Methods for Analyzing Models of Rod-Shaped Bacteria

### 2025-07-26 paper/main.tex
write visualization part

% --------------------------------------------------------------------------------------------------
\subsection{Visualization (TODO)}
\label{section:generating-masks-and-images}
\begin{itemize}
    \item use pyvista (VTK) for Visualization
    \item always do 3D render. This way, we can have overalp between agents and produce more
        realistic masks.
    \item Assign unique color values for idividual agents.
        We have a 1:1 mapping between agent identifier and color! (for each simulation different
        though)
    \item use parallel projection; this should be most close to what microscopes measure due to the
        relatively large distance between object and lens.
    \item use some noise to smear out the rods; this is already similar to microscopic images; use
        more like this and deal with real-world defects to make it more realistic (future, not now)
    \item Figure~\ref{fig:progression-image-generation} shows steps in generating image.
    \item Reflectivity, placement of light etc. can already introduce some heterogeneity in the
        visual of the rod.
    \item Plan to implement noise on the mesh itself; make it have rough surface (like real rods as
        seen under electron microscopy)
\end{itemize}

\paragraph{Algorithm}
\begin{enumerate}
    \item Get position of cell
    \item Insert spheres at each vertex
    \item Insert cylinder along edges
    \item Combine individual objects into one big mesh
    \item Use \ac{pbr}
\end{enumerate}

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.3\textwidth]
        {docs/source/_static/09395645494836445480/raw_pv/000000400.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.3\textwidth]
        {docs/source/_static/09395645494836445480/images/000000400.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.3\textwidth]
        {docs/source/_static/09395645494836445480/masks/000000400.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}
    \caption{
        (A) shows the raw result from combining sphere and cylinder meshes to obtain the shape of a
        bacterial rod.
        This render contains lighting coming from 2 light sources which thus produces a small glow
        on the upper part of the meshes.
        Subfigure (B) takes this image and applies noise to it in order to make it look more
        like an actual microscopic image.
        (C) The render is projected along the z-axis and flat numerical values are assigned to the
        bacteria.
        The resuling render can have overlaps in which case the color of the rod is displayed that
        is higher along the z-direction.
    }
    \label{fig:progression-image-generation}
\end{figure}

>>>>

% --------------------------------------------------------------------------------------------------
\subsection{Visualization}
\label{section:generating-masks-and-images}

For visualization we use pyvista, which provides high-level wrappers around the popular VTK
library~\cite{vtkBook,Sullivan2019}.
The first stage of our visualization is to generate 3D meshes for each bacterium, which we consider
an intermediate representation.
We insert spheres of a given radius $r$ at each vertex $\vec{v}_i$ and connect adjacent vertices
$\vec{v}_i,\vec{v}_{i+1}$ with a cylinder, also of radius $r$ which is positioned such that its core
is identical to the connecting line segment between the two vertices.
Finally, we combine the meshes of all of these objects to obtain a single mesh of one bacterium.
A render of this result can be seen in figure~\ref{fig:progression-image-generation}.
Since we use a 3D rendering engine as an intermediate representation, the resulting image can have
overlaps of bacteria which are placed at higher or lower values along the z-direction.

To generate masks, we assign unique colors to every mesh of the respective cellular agent.
This mapping is unique, ensuring no two agents will use identical colors.
We then use a parallel projection and remove any existing light sources in order to obtain a solid
color for every cell-agent.
The resulting image is shown in subfigure (B).
Subfigure (C) shows an example of a simplified way to generate microscopic images.
The properties of the meshes such as roughness and reflectivity have been adjusted and we use
physically-based rendering of VTK.
The result of this render is augmented by introducing white noise on the pixel values at each point.
This approach is very rudimentary and we hope to extend these methods in the future.
In order to properly generate data, we need to deal with microscopic imaging techniques and possible
defects that can occurr.
It is crucial to get this part of the image-generation process right in order to obtain
near-realistic images that can be further reused.

% \begin{itemize}
%     \item use pyvista (VTK) for Visualization
%     \item always do 3D render. This way, we can have overalp between agents and produce more
%         realistic masks.
%     \item Assign unique color values for idividual agents.
%         We have a 1:1 mapping between agent identifier and color! (for each simulation different
%         though)
%     \item use parallel projection; this should be most close to what microscopes measure due to the
%         relatively large distance between object and lens.
%     \item use some noise to smear out the rods; this is already similar to microscopic images; use
%         more like this and deal with real-world defects to make it more realistic (future, not now)
%     \item Figure~\ref{fig:progression-image-generation} shows steps in generating image.
%     \item Reflectivity, placement of light etc. can already introduce some heterogeneity in the
%         visual of the rod.
%     \item Plan to implement noise on the mesh itself; make it have rough surface (like real rods as
%         seen under electron microscopy)
% \end{itemize}
% 
% \paragraph{Algorithm}
% \begin{enumerate}
%     \item Get position of cell
%     \item Insert spheres at each vertex
%     \item Insert cylinder along edges
%     \item Combine individual objects into one big mesh
%     \item Use \ac{pbr}
% \end{enumerate}

\begin{figure}[H]
    \centering
    % \begin{tikzonimage}[width=0.9\textwidth]{visualization.pdf}
    %     \node at (0.025, 0.975)[anchor=north west, rectangle, draw, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    % \end{tikzonimage}
    \begin{tikzonimage}[width=0.3\textwidth]
        {docs/source/_static/09395645494836445480/raw_pv/000000400.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.3\textwidth]
        {docs/source/_static/09395645494836445480/masks/000000400.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.3\textwidth]
        {docs/source/_static/09395645494836445480/images/000000400.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}
    \caption{
        (A) shows the result from combining sphere and cylinder meshes to obtain the shape of a
        bacterial rod.
        This render contains lighting coming from 2 light sources which thus produces a small glow
        on the upper part of the meshes.
        (B) Unique color values are assigned to each agent and light sources are removed.
        Afterwards, the image is rendered by projecting along the z-axis.
        (C) We use physically-based rendering and assign properties such as roughness and
        reflectivity to generate an initial render.
        Afterwards, we apply white noise to the pixel values of image.
        In the future, we hope to extend this approach in order to construct near-realistic
        microscopic images from synthetic data.
    }
    \label{fig:progression-image-generation}
\end{figure}
